\bf Model Visualization Techniques for a Social Network Model

### 2017-08-19 paper.Rnw
up to section 4.1

Because we require binary edges for our models, we focus only on very strong collaborations. The networks we constructed for the four senates during President Obama's administration are shown in Figure~\ref{fig:senateplot} so for our collaboration networks, $x$, edges are defined as 
\[
  x_{ij} =
  \begin{cases}
                                   1 &  \text{if } WPC_{ij} > 0.25 \\
                                   0 & \text{if } WPC_{in \leq 0.25}.
  \end{cases}
\]

}
\section{Model Visualizations} \label{sec:ModelVis}

The concept of model visualization was developed to complement traditional model diagnostic tools. Typically, numerical summaries such as $R^2$ are used to assess model fit, and visualizations, like residual plots, are restricted to the data. Every good data analysis includes both numerical and visual summaries of the data, so why should model description be restricted to numerical summaries? \citeauthor{modelvis} outline the three different parts of a model: the model family, the model form, and the fitted model. The latter is primarily what one thinks of first when considering a model in a data analysis, where a model is fit to data, and parameter estimates are reported. In the context of SAOMs, the fitted model contains the estimated rate parameters and the estimated objective function parameters. The model form describes the the model before the fitting process and how variables describe the data in the model. In SAOMs, the model form includes description of the rate and objective functions and the variables therein that describe how the network evolves over time. Finally, the model family is the most broad description of the model. This is the type of model that you wish to fit to the data, and is chosen based on the problem and knowledge at had. For example, we chose to use a SAOM to model network data over an exponential random graph model (ERGM) because we believe that actor-level variables effect network dynamics. These three parts of a "model" are should be visualized according to the three principals of model visualization: display the model in the data space, look at a collection of models, and explore the process of fitting the model, not just the end result.

>>>>

Because we require binary edges for our models, we focus only on very strong collaborations. For our senate collaboration networks, $x$, edges are defined as 
\[
  x_{ij} =
  \begin{cases}
                                   1 &  \text{if } WPC_{ij} > 0.25 \\
                                   0 & \text{if } WPC_{ij} \leq 0.25.
  \end{cases}
\]
The networks we constructed for the four senates during President Obama's administration are shown in Figure~\ref{fig:senateplot}. 
}
\section{Model Visualizations} \label{sec:ModelVis}

Every good data analysis includes both numerical and visual summaries of the data, so why restrict model description and diagnostics to numerical summaries? The concept of model visualization was developed to complement traditional model diagnostic tools. Typically, numerical summaries such as $R^2$ are used to assess model fit, and the occasional visualization, like a residual plot, are used to determine how well the model fits the data. \citeauthor{modelvis} outline three separate ideas, each of which are can be referred to simply as the "model": the model family, the model form, and the fitted model. The latter is primarily what one thinks of first when considering a model in a data analysis, where specified a model is fit to data, and parameter estimates and other numerical summaries, such as $R^2$ are reported. In the context of SAOMs, the fitted model contains the form of the rate and objective functions, the estimated rate parameters, and the estimated objective function parameters. The model form describes the the model \textit{before} the fitting process, defining which parameters are in the model within the context of the larger model family. In SAOMs, the model form includes description of the rate and objective functions and the variables therein that describe how the network evolves over time. Finally, the model family is the broadest description of the model. This is the type of model that you wish to fit to the data, and is chosen based on the problem, data, and knowledge at hand. For example, we chose to use a SAOM to model network data over an exponential random graph model (ERGM) because we believe that actor-level variables effect network structure and formation, and we wanted to model the network changes over time. These three uses of the term ``model" can each be visualized according to the three principals of model visualization: display the model in the data space, look at a collection of models, and explore the process of fitting the model, not just the end result. We turn our focus to the fitted model and the model form. Specifically, we want to know how the model form affects the fitted model. Using our example data sets and our visualization toolbox, we hope to answer this question for SAOMs.
